---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
###1. Load the data
```{r}
setwd("U:/General/Data_Scientist_Course/Course5_ReproducibleResearch/Project1/RepData_PeerAssessment1")
file<-"activity.zip"
unzip(file,unzip = "internal")
data<-read.csv("activity.csv")
```
###2. Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
data$date<-as.POSIXct(data$date,"%Y-%m-%d",tz="GMT")
```

## What is mean total number of steps taken per day?
###1. Calculate the total number of steps taken per day
```{r}
TotalPerDay<-tapply(data$steps,data$date,sum)
```
###2. Make a histogram of the total number of steps taken each day
```{r}
hist(TotalPerDay)
```

###3. Calculate and report the mean and median of the total number of steps taken per day
```{r}
#Mean Steps/Day
mean<-mean(TotalPerDay,na.rm=TRUE)
#Median Steps/Day
median<-median(TotalPerDay,na.rm=TRUE)
```
The mean number of steps taken was `r mean`. The median number of steps taken was `r median`.

## What is the average daily activity pattern?
###1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
#Get Average Number of Steps per Interval
AvgPerInterval<-tapply(data$steps,data$interval,mean,na.rm=TRUE)
#plot them
plot(AvgPerInterval,type="l")
```

###2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
#Calculate Max
max<-max(AvgPerInterval)
#Get index of interval with max
index<-match(max,AvgPerInterval)
#Pull corresponding time interval
MaxInterval<-unique(data$interval)[index]
```
The 5-minute interval with the max number of step on average was `r MaxInterval`.

## Imputing missing values
###1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
NumberNAs<-sum(is.na(data$steps))
```
The total number of missing values in this dataset is `r NumberNAs`.

###2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
###3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
#fill missing and create new
library(dplyr)
newdata<-data %>% group_by(interval) %>% mutate(steps = ifelse(is.na(steps),mean(steps, na.rm=TRUE),steps))
```

###4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r}
newTotalPerDay<-tapply(newdata$steps,data$date,sum)
hist(newTotalPerDay)
```

###5. Calculate and report the mean and median of the total number of steps taken per day
```{r}
#Mean Steps/Day
mean<-mean(newTotalPerDay)
#Median Steps/Day
median<-median(newTotalPerDay)
```
The new mean number of steps taken was `r mean`. The new median number of steps taken was `r median`.
Now the mean and median are exactly the same and both are higher than before.

## Are there differences in activity patterns between weekdays and weekends?
###1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
newdata$daytype<-ifelse(weekdays(newdata$date) %in% c("Saturday","Sunday"),"weekend","weekday")
```
###2.Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r}
#Average data by day type and interval
library(plyr)
avgdata<-ddply(newdata, .(daytype, interval), summarize, avgsteps = mean(steps))
library(ggplot2)
ggplot(avgdata, aes(x = interval, y = avgsteps)) + geom_line() + facet_wrap(~daytype)
```